---
title: "Exercise5"
author: Chutian Chen cc4515; Congcheng Yan cy2550; Mingrui Liu ml4404
output: pdf_document
---

## (3)
Let X be Poisson Distribution.
$EY=\frac{1}{1-e^{-\mu}-\mu e^{-\mu}}(EX-P(X=0)*0-P(X=1)*1)$
$=\frac{\mu-\mu e^{-\mu}}{1-e^{-\mu}-\mu e^{-\mu}}$

## (4)
Log likelihood: $\sum_{i=1}^{n}(-\mu+y_iln(\mu)-ln(y_i!)-ln(1-e^{-\mu}-\mu e^{-\mu}))$

```{r}
library(ggplot2)
freq <- c(179,51,17,6,8,1,0,2)
y <- seq(2,9,1)
data <- as.data.frame(y)
data$freq <- freq
log_mle <- function(miu) {
  likehood <- function(miu,y) {
    return(-miu+y*log(miu)-log(factorial(y))-log(1-exp(-miu)-miu*exp(-miu)))
  }
  return(sum(likehood(miu,data$y)*data$freq))
}
miu <- seq(0.8,1.6,0.1)
mle <- sapply(miu,log_mle)
ggplot() + geom_line(aes(x = miu, y = mle))
```

## (5)
```{r}
library(optim.functions)
initial <- c(1)
optim(initial, a <- function(x){return(-log_mle(x))}, method = "BFGS")
```

We use BFGS. The MLE of $\mu$ is 1.398.

## (6)

$\sqrt{n}(\hat\mu_{MLE}-\mu)\rightarrow^D N(0,I^{-1}(\mu))$

$I(\mu)=-E(\frac{d^2(-\mu+yln(\mu)-ln(y!)-ln(1-e^{-\mu}-\mu e^{-\mu}))}{d\mu^2})$

$=E(\frac{y}{\mu^2}+\frac{e^{\mu}-\mu e^{\mu}-1}{(e^{\mu}-1-\mu)^2})=\frac{(e^{\mu}-1)^2-\mu^2 e^{\mu}}{\mu(e^{\mu}-1-\mu)^2}$

## (7)
Through the asymptotic distribution of $\mu_{MLE}-\mu$, we can get a 0.95 asymptotic confidence interval.

$[\hat\mu_{MLE}-\frac{Z_{0.975}}{\sqrt{nI(\hat\mu)}},\hat\mu_{MLE}+\frac{Z_{0.975}}{\sqrt{nI(\hat\mu)}}]$

```{r}
m <- 1.398
i <- ((exp(m)-1)^2-m^2*exp(m))/m/(exp(m)-1-m)^2
interval <- qnorm(0.975)/sqrt(sum(data$freq)*i)
print(c(m-interval,m+interval))
```
95% CI is [1.20,1.60]

## (8)

