---
title: "Homework 3 Exercise 2"
author: "Chutian Chen cc4515; Congcheng Yan cy2550; Mingrui Liu ml4404"
date: "2020/4/16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
```{r message=FALSE}
library(mgcv)
cars$speed2 <- (cars$speed)^2
model1 = lm(dist~speed + speed2, data = cars)
summary(model1)
AIC(model1)
```

```{r}
model2 = lm(dist~speed, data = cars)
summary(model2)
AIC(model2)
```
From the results, the model with speed2 is a better fit for the data.

## Question 2
From the hypothesis test result in Question 1, the p values are not so good. We can take the reaction time instead. Due to the different reaction time, it could lead to a non-linear effection to the model. Also, we can test the combination of the speed and the reaction time to see their influence to the regression.



## Question 3

```{r}

myf <- function (X, y) {
  qrx <- qr(X) ## returns a QR decomposition object
  Q <- qr.Q(qrx,complete=TRUE) ## extract Q
  R <- qr.R(qrx) ## extract R
  f <- t(Q)%*%y
  beta <- backsolve(R,f)
  return(beta)
}
```
## Question 4

```{r}
test1 <- myf(model.matrix(dist ~ speed + speed2, cars), data.matrix(cars[,c("dist")]))
test1
```

The coefficients are the same as Question 1, so it works.

## Question 5

```{r}
myf2 <- function (X, y) {
  qrx <- qr(X) ## returns a QR decomposition object
  Q <- qr.Q(qrx,complete=TRUE) ## extract Q
  R <- qr.R(qrx) ## extract R
  f <- t(Q)%*%y
  beta <- backsolve(R,f)
  
  res <- y-X%*%beta
  s <- as.vector(t(res)%*%res/(nrow(X)-ncol(X)))
  var <- solve(R)%*%t(solve(R))*s
  return(list(beta=beta, std=sqrt(as.matrix(diag(var),ncol=ncol(X)))))
}
```
```{r}
test2 <- myf2(model.matrix(dist ~ speed + speed2, cars), data.matrix(cars[,c("dist")]))
test2$std
```
## Question 6


```{r}
mat= model.matrix(dist ~ speed + speed2, cars)
pt(test2$beta / test2$std, nrow(mat)-ncol(mat), lower.tail=FALSE)
```


















