---
title: "Homework 3 Exercise 1"
author: "Chutian Chen cc4515; Congcheng Yan cy2550; Mingrui Liu ml4404"
date: "2020/4/16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r}
library(dplyr)
milk <- read.csv("milk.csv", sep = ",", header = TRUE, 
                 col.names = c("Month", "Production"))
```

```{r}
library(lubridate)
milk <- milk %>% 
  mutate(Month=ymd(Month, truncated = 1)) %>% 
  mutate(Row=row_number())
```


```{r}
library(ggplot2)
ggplot(milk, aes(x=Month, y=Production)) + 
  geom_line() + 
  geom_smooth(method = "lm") +
  theme_bw()
```
```{r}
milk.lm <- lm(Production ~ Row, milk)
milk.lm
```
```{r}
res = resid(milk.lm)
plot(milk$Row, res, 
    ylab="Residuals", xlab="Row")
abline(0, 0)
```

We make the regression with index instead of the timestamp. We can find a good inceasing trend through this linear regression. Also, the residuals seem to have a random value with mean = 0.


## Question 2

```{r, message=FALSE}
library(forecast)
ggAcf(milk.lm$residuals)
```

```{r}
ggPacf(milk.lm$residuals)
```

We can find a obvious seasonal trend in the graph. Maybe AR(1), AR(2) will have a better result.

## Question 3
```{r}
AR1 <- Arima(res %>% diff(12), order=c(1,0,0))
AR1
```
```{r}
checkresiduals(AR1)
```
```{r}
AR2 <- Arima(res %>% diff(12), order=c(2,0,0))
AR2
```
```{r}
checkresiduals(AR2)
```
```{r}
MA1 <- Arima(res %>% diff(12), order=c(0,0,1))
MA1
```
```{r}
checkresiduals(MA1)
```
```{r}
MA2 <- Arima(res %>% diff(12), order=c(0,0,2))
MA2
```
```{r}
checkresiduals(MA2)
```

If we pick AIC as the evaluation, AR1 and AR2 are similar. AR2 is the best fit among the 4 fits.



## Question 4

```{r}
AIC(Arima(res %>% diff(12), order=c(2,0,1)))
AIC(Arima(res %>% diff(12), order=c(2,0,2)))
AIC(Arima(res %>% diff(12), order=c(2,0,3)))
AIC(Arima(res %>% diff(12), order=c(3,0,1)))
AIC(Arima(res %>% diff(12), order=c(3,0,2)))
AIC(Arima(res %>% diff(12), order=c(3,0,3)))
```
So, the top 2 fits are ARMA(2,3) and ARMA(3,1)













