---
title: "Exercise3"
author: "Chutian Chen cc4515; Congcheng Yan cy2550; Mingrui Liu ml4404"
date: "3/1/2020"
output: pdf_document
---

```{r message=FALSE}
library(dplyr)
library(lubridate)
```

## (1)

Let $X_i$ represent whether the ith day is rainy.

$a_1$: $P(X_i=0|X_{i-1}=0)$

$a_2$: $P(X_i=1|X_{i-1}=0)$

$a_3$: $P(X_i=0|X_{i-1}=1)$

$a_4$: $P(X_i=1|X_{i-1}=1)$

## (2)

$P(X_n=0)=P(X_n=0|X_{n-1}=0)P(X_{n-1}=0)+P(X_n=0|X_{n-1}=1)P(X_{n-1}=1) \\ = a_1P(X_n=0)+a_3(1-P(X_n=0))$

So $P(X=0)=\frac{a_3}{1-a_1+a_3}$

## (3)

```{r}
data <- read.csv('./CentralPark.csv', header = T)
data$DATE <- as.POSIXct(strptime(as.character(data$DATE), "%m/%d/%y"))
data <- data %>% mutate(rain = if_else(PRCP>=1.5,TRUE,FALSE))
data$month <- month(data$DATE)
data$rain_tomo <- append(data$rain,c(NA))[2:(length(data$rain)+1)]
```
```{r}
rain_day <- nrow(data %>% filter(month == 7, rain))
nrain_day <- nrow(data %>% filter(month == 7, !rain))
r_r <- nrow(data %>% filter(month == 7, rain, rain_tomo))
r_nr <- nrow(data %>% filter(month == 7, rain, !rain_tomo))
nr_r <- nrow(data %>% filter(month == 7, !rain, rain_tomo))
nr_nr <- nrow(data %>% filter(month == 7, !rain, !rain_tomo))
a1 <- r_r/rain_day
a3 <- nr_r/nrain_day
a2 <- r_nr/rain_day
a4 <- nr_nr/nrain_day
print(c(a1,a2,a3,a4))
```

## (4)

$H_0:a_0=a_4$, $H_1:a_0\neq a_4$

We can use $p_{00}$ and $p_{11}$ to represent $a_0$ and $a_4$.

$\hat p_{00}\rightarrow^DN(\hat p_{00},\frac{\hat p_{00}\left(1-\hat p_{00}\right)}{n_0})$

$\hat p_{11}\rightarrow^DN(\hat p_{11},\frac{\hat p_{11}\left(1-\hat p_{11}\right)}{n_1})$

Because $p_{00}$ and $p_{11}$ are asymptotic independent, under $H_0$

$\hat p_{00}-\hat p_{11}\rightarrow^DN(0,\frac{\hat p_{00}\left(1-\hat p_{00}\right)}{n_0}-\frac{\hat p_{11}\left(1-\hat p_{11}\right)}{n_1})$

```{r}
pnorm((a1-a4)/sqrt(a1*(1-a1)/rain_day+a4*(1-a4)/nrain_day))
```

We can reject $H_0$.

## (5)
```{r}
data$rain_tomo2 <- append(data$rain_tomo,c(NA))[2:(length(data$rain_tomo)+1)]

r_r_r <- nrow(data %>% filter(month == 7, rain, rain_tomo, rain_tomo2))
r_r_nr <- nrow(data %>% filter(month == 7, rain, rain_tomo, !rain_tomo2))
r_nr_r <- nrow(data %>% filter(month == 7, rain, !rain_tomo, rain_tomo2))
r_nr_nr <- nrow(data %>% filter(month == 7, rain, !rain_tomo, !rain_tomo2))
nr_r_r <- nrow(data %>% filter(month == 7, !rain, rain_tomo, rain_tomo2))
nr_r_nr <- nrow(data %>% filter(month == 7, !rain, rain_tomo, !rain_tomo2))
nr_nr_r <- nrow(data %>% filter(month == 7, !rain, !rain_tomo, rain_tomo2))
nr_nr_nr <- nrow(data %>% filter(month == 7, !rain, !rain_tomo, !rain_tomo2))

p000 <- r_r_r / (r_r_r + r_r_nr)
p001 <- r_r_nr / (r_r_r + r_r_nr)
p010 <- r_nr_r / (r_nr_r + r_nr_nr)
p011 <- r_nr_nr / (r_nr_r + r_nr_nr)
p100 <- nr_r_r / (nr_r_r + nr_r_nr)
p101 <- nr_r_nr / (nr_r_r + nr_r_nr)
p110 <- nr_nr_r / (nr_nr_r + nr_nr_nr)
p111 <- nr_nr_nr / (nr_nr_r + nr_nr_nr)

p00 <- a1
p01 <- a2
p10 <- a3
p11 <- a4
```

$H_0$: the first order model holds.

$H_1$: the first order model doesn't hold.

Use likelihood ratio test:

$\begin{aligned}\Lambda_{n} &=2\left\{\ell(\hat{\mathbf{P}})_{\text {second order}}-\ell(\hat{\mathbf{P}})_{\text {first order}}\right\}=2\left\{\sum_{r=1}^{S} \sum_{s=1}^{S} \sum_{t=1}^{S} n_{r s t} \log \hat{p}_{r s t}-\sum_{s=1}^{S} \sum_{t=1}^{S}n_{. st} \log \hat{p}_{st}\right\} \\&=2\left\{\sum_{r=1}^{S} \sum_{s=1}^{S} \sum_{t=1}^{S} n_{r s t} \log \hat{p}_{r s t}-\sum_{r=1}^{S} \sum_{s=1}^{S} \sum_{t=1}^{S} n_{r s t} \log \hat{p}_{st}\right\}=2 \sum_{r=1}^{S} \sum_{s=1}^{S} \sum_{t=1}^{S} n_{r s t} \log \left(\frac{\hat{p}_{r s t}}{\hat{p}_{s t}}\right)\end{aligned}$

$\Lambda_{n} \frac{\mathcal{D}}{n \rightarrow \infty} \chi_{(S-1)^{2}}^{2}$ 

```{r}
lambda <- (r_r_r*log(p000/p00) + r_r_nr*log(p001/p01) + r_nr_r*log(p010/p10) + r_nr_nr*log(p011/p11) + nr_r_r*log(p100/p00) + nr_r_nr*log(p101/p01) + nr_nr_r*log(p110/p10) + nr_nr_nr*log(p111/p11)) * 2
pchisq(lambda,2)
```

We can accept $H_0$. Higher order chain doesn't improve the fit of the data.



