---
title: "Exercise5"
author: "Chutian Chen cc4515; Congcheng Yan cy2550; Mingrui Liu ml4404"
date: "3/1/2020"
output: pdf_document
---

```{r}
transplant2 <- read.table("./transplant.txt", quote="\"", comment.char="")
colnames(transplant2)<-c("time","type","s")
```

## (1)

Because the failure rate doesn't dependent on the number of individuals.

Let the probability of individuals that die/relapse be $p_A^{(k)}=p_B^{(k)}=p$ (under $H_0$).

So $P(y^{(k)}=m,n_d^{(k)}=m_d|n_A^{(k)}=m_A,n_B^{(k)}=m_B)={m_A\choose m}p^m(1-p)^{m_A-m}{m_B\choose m_d-m}p^{m_d-m}(1-p)^{m_B-m_d+m}={m_A\choose m}{m_B\choose m_d-m}p^{m_d}(1-p)^{m_A+m_B-m_d}$

$P(n_d^{(k)}=m_d|n_A^{(k)}=m_A,n_B^{(k)}=m_B)={m_A+m_B\choose m_d}p^{m_d}(1-p)^{m_A+m_B-m_d}$

So $P(y^{(k)}=m|n_d^{(k)}=m_d,n_A^{(k)}=m_A,n_B^{(k)}=m_B) \\ =\frac{P(y^{(k)}=m,n_d^{(k)}=m_d|n_A^{(k)}=m_A,n_B^{(k)}=m_B)}{P(n_d^{(k)}=m_d|n_A^{(k)}=m_A,n_B^{(k)}=m_B)}=\frac{{m_A\choose m}{m_B\choose m_d-m}}{{m_A+m_B\choose m_d}}$

It's $\text {HyperGeometric}\left(n_{A}^{(k)}+n_{B}^{(k)}, n_{A}^{(k)}, n_{d}^{(k)}\right)$.

## (2)

$P=\frac{{n_A\choose y}{n_B\choose n_d-y}}{{n\choose n_d}}$

Let $Y_{i}=\left\{\begin{array}{ll} 1 & \text { if the } i \text { th selection is under treatment A } \\ 0 & \text { otherwise } \end{array}\right.$

So $P(Y_i=1)=\frac{n_A}{n}$, $P(Y_i=1,Y_j=1)=\frac{n_A}{n}\frac{n_A-1}{n}$

$E^{(k)}=E(\sum Y_i)=n_dEY=\frac{n_{A}^{(k)} n_{d}^{(k)}}{n^{(k)}}$

$Var^{(k)}=Var(\sum Y_i)=n_dVarY+n_d(n_d-1)Cov(Y_i,Y_j)$

$VarY=\frac{n_An_B}{n^2}$, $Cov(Y_i,Y_j)=EY_iY_j-EY_iEY_j=P(Y_i=1,Y_j=1)-(\frac{n_A}{n})^2=-\frac{n_An_B}{n^2(n-1)}$

Then $Var^{(k)}=\frac{n_An_Bn_d}{n^2}-\frac{n_An_Bn_d(n_d-1)}{n^2(n-1)}=\frac{n_{A}^{(k)} n_{B}^{(k)} n_{d}^{(k)} n_{s}^{(k)}}{\left(n^{(k)}\right)^{2}\left(n^{(k)}-1\right)}$

## (3)

Let condition $\left(n_{A}^{(k)}, n_{B}^{(k)}, n_{d}^{(k)}\right)$ be $C_i$.

$\operatorname{Var}\left[y^{(k)}-E^{(k)}\right]=\operatorname{Var}[\mathbb{E}[y^{(k)}-E^{(k)}| C_i]]+\mathbb{E}[\operatorname{Var}[y^{(k)}-E^{(k)}| C_i]]$

As we showed in (2), $\mathbb{E}[y^{(k)}-E^{(k)}| C_i]=0$, $\operatorname{Var}[y^{(k)}-E^{(k)}| C_i]=V^{(k)}$.

So $\operatorname{Var}\left[y^{(k)}-E^{(k)}\right]=E[V^{(k)}]$

## (4)

$\operatorname{Var}\left[\sum_{k=1}^{K}\left(y^{(k)}-E^{(k)}\right)\right]=\sum_{k=1}^{K} \operatorname{Var}\left[y^{(k)}-E^{(k)}\right]+2\sum_{i=1}^{K-1}\sum_{j=i+1}^{K}\operatorname{Cov}\left[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}\right]$

Let the condition $\left(n_{A}^{(i)}, n_{B}^{(i)}, n_{d}^{(i)},n_{A}^{(j)}, n_{B}^{(j)}, n_{d}^{(j)}\right)$ be $C_{ij}$

With Law of Total Variance, we can get

$\operatorname{Cov}\left[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}\right] \\ =\operatorname{Cov}[\mathbb{E}[y^{(i)}-E^{(i)}| C_{ij}],\mathbb{E}[y^{(j)}-E^{(j)}| C_{ij}]]+\mathbb{E}[\operatorname{Cov}[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}| C_{ij}]]$

$\mathbb{E}[y^{(i)}-E^{(i)}| C_{ij}]=\mathbb{E}[y^{(i)}-E^{(i)}| C_i]=0$

So $\operatorname{Cov}[\mathbb{E}[y^{(i)}-E^{(i)}| C_{ij}],\mathbb{E}[y^{(j)}-E^{(j)}| C_{ij}]]=Cov(0,0)=0$

$\operatorname{Cov}[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}| C_{ij}]=\operatorname{Cov}[y^{(i)},y^{(j)}| C_{ij}]$

Under the condition $C_{ij}$, $y^{(i)},y^{(j)}$ are two independent hyperGeometric distributions, so the covariance of them is 0.

Then $\operatorname{Cov}\left[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}\right]=0 \;\;\forall i\neq j$

So $\operatorname{Var}\left[\sum_{k=1}^{K}\left(y^{(k)}-E^{(k)}\right)\right]=\sum_{k=1}^{K} \operatorname{Var}\left[y^{(k)}-E^{(k)}\right]=\sum_{k=1}^{K} \mathbb{E}\left[V^{(k)}\right]$

## (5)

```{r message=FALSE}
library(dplyr)
```

```{r}
n <- nrow(transplant2)
k <- nrow(transplant2 %>% filter(s==1))
na <- nrow(transplant2 %>% filter(type==1))
nb <- nrow(transplant2 %>% filter(type==2))
df <- transplant2[order(transplant2$time),]
df <- df %>% filter(s==1)
y = c()
e = c()
v = c()
for (i in 1:k) {
  
  e[i] <- na/(na+nb)
  v[i] <- na*nb*(na+nb-1)/(na+nb)^2/(na+nb-1)
  
  if (df$type[i]==1) {
    y[i] <- 1
    na <- na - 1
  } else {
    y[i] <- 0
    nb <- nb - 1
  }
}

z <- sum(y-e)/sqrt(sum(v))
print(pnorm(z))
```
We can accept $H_0$ at 0.05 significant level.




